---
title: "Getting Started with fmapsR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with fmapsR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`fmapsR` provides a high-level functional maps interface that is not restricted to shape meshes. The same API is used for:

1. Pairwise functional map estimation.
2. Multi-dataset synchronization and consensus maps.
3. Generic multidimensional datasets (time-series, tensors, tabular embeddings).

## Core API surface

```{r}
library(fmapsR)

c(
  "fm_domain()", "fm_basis()", "fm_match()", "fm_refine()",
  "fm_network()", "fm_sync()", "fm_consensus()", "fm_transfer()"
)
```

## Vignettes in this package

1. `pairwise-functional-maps`: concise pairwise map workflow.
2. `multi-dataset-sync-consensus`: synchronization + consensus across multiple datasets.
3. `non-shape-multidimensional-data`: functional maps on generic multidimensional data.

## Package metadata

```{r}
fm_package_info()
```
