---
title: "Multi-Dataset Synchronization and Consensus"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multi-Dataset Synchronization and Consensus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(fmapsR)
set.seed(42)

make_domain <- function(seed, n = 20, k = 6) {
  set.seed(seed)
  latent <- matrix(rnorm(n * 4), nrow = n, ncol = 4)
  fm_basis(
    fm_domain_generic(n_samples = n, data = latent, operator = diag(seq_len(n))),
    k = k,
    solver = "base",
    cache = FALSE
  )
}

d1 <- make_domain(101)
d2 <- make_domain(102)
d3 <- make_domain(103)

desc1 <- d1$data
desc2 <- d1$data + matrix(rnorm(length(d1$data), sd = 0.03), nrow = nrow(d1$data))
desc3 <- d1$data + matrix(rnorm(length(d1$data), sd = 0.04), nrow = nrow(d1$data))
```

## End-to-end network sync workflow

The core workflow below uses **10 lines** of API calls.

```{r}
fit12 <- fm_refine(fm_match(d1, d2, descriptors = list(source = desc1, target = desc2), optimizer = "cg", cg_maxit = 3), method = "icp", nit = 2)
fit23 <- fm_refine(fm_match(d2, d3, descriptors = list(source = desc2, target = desc3), optimizer = "cg", cg_maxit = 3), method = "icp", nit = 2)
fit31 <- fm_refine(fm_match(d3, d1, descriptors = list(source = desc3, target = desc1), optimizer = "cg", cg_maxit = 3), method = "icp", nit = 2)
net <- fm_network(list(a = d1, b = d2, c = d3), directed = TRUE)
net <- fm_network_add_map(net, "a", "b", fit12)
net <- fm_network_add_map(net, "b", "c", fit23)
net <- fm_network_add_map(net, "c", "a", fit31)
synced <- fm_sync(net, mode = "cycle", nit = 10)
consensus <- fm_consensus(synced, method = "latent_clb")
fm_network_report(net, consensus$network)
```

## Notes

- `fm_sync()` reduces cycle inconsistencies across maps.
- `fm_consensus(method = "latent_clb")` refines synchronized maps with latent-basis optimization.
- The same pattern scales to larger graph topologies.

## Batch Controls for Larger Collections

For bigger runs, use `fm_network_match()` to batch edges and stream descriptors:

```{r, eval = FALSE}
net2 <- fm_network_match(
  domains = list(a = d1, b = d2, c = d3),
  descriptors = list(a = desc1, b = desc2, c = desc3),
  edges = c("a->b", "b->c", "c->a"),
  pair_batch_size = 2,
  descriptor_batch_size = 8,
  optimizer = "cg"
)
```
