---
title: Multi-Dataset Synchronization and Consensus
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2.0
    css: albers.css
    includes:
      in_header: albers-header.html
params:
  family: teal
  preset: homage
resource_files:
- albers.css
- albers.js
- albers-header.html

vignette: |
  %\VignetteIndexEntry{Multi-Dataset Synchronization and Consensus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(fmapsR)
set.seed(42)

make_domain <- function(n = 8, k = 4) {
  op <- diag(seq_len(n))
  fm_basis(
    fm_domain_generic(n_samples = n, operator = op),
    k = k,
    solver = "base",
    cache = FALSE
  )
}

random_orth <- function(k) {
  qr.Q(qr(matrix(rnorm(k * k), nrow = k, ncol = k)))
}

make_noisy_cycle_network <- function(k = 4, noise_sd = 0.35) {
  d1 <- make_domain(k = k)
  d2 <- make_domain(k = k)
  d3 <- make_domain(k = k)

  X1 <- diag(k)
  X2 <- random_orth(k)
  X3 <- random_orth(k)

  C12 <- X2 %*% t(X1) + matrix(rnorm(k * k, sd = noise_sd), nrow = k)
  C23 <- X3 %*% t(X2) + matrix(rnorm(k * k, sd = noise_sd), nrow = k)
  C31 <- X1 %*% t(X3) + matrix(rnorm(k * k, sd = noise_sd), nrow = k)

  net <- fm_network(list(a = d1, b = d2, c = d3), directed = TRUE)
  net <- fm_network_add_map(net, "a", "b", C12)
  net <- fm_network_add_map(net, "b", "c", C23)
  net <- fm_network_add_map(net, "c", "a", C31)
  net
}

net <- make_noisy_cycle_network()
```

## End-to-end network sync workflow

This vignette starts after pairwise edge maps already exist. The synthetic maps below are intentionally inconsistent, so the synchronization report has something real to show.

```{r}
synced <- fm_sync(net, mode = "cycle", nit = 10)
sync_report <- fm_network_report(synced)
consensus <- fm_consensus(synced, method = "latent_clb", latent_nit = 8)

sync_report

c(
  pre_cycle_median = synced$diagnostics$pre_cycle_median,
  post_cycle_median = synced$diagnostics$post_cycle_median,
  cycle_reduction_ratio = synced$diagnostics$cycle_reduction_ratio,
  consensus_cycle_median = consensus$diagnostics$consensus_cycle_median,
  consensus_confidence_median = consensus$uncertainty$summary$median
)
```

## Notes

- `fm_sync()` is doing visible work here because the pre-sync cycle median starts above zero and the report compares it to the synchronized network.
- `fm_consensus(method = "latent_clb")` keeps the network in a low-cycle-consistency regime and adds an uncertainty summary for the consensus edges.
- If your inputs are raw descriptors rather than pre-fit maps, use `fm_network_match()` to build the initial network first.

## Build the Edge Maps from Raw Descriptors

For bigger runs, `fm_network_match()` batches the pairwise fits before synchronization:

```{r, eval = FALSE}
d1 <- make_domain()
d2 <- make_domain()
d3 <- make_domain()

desc1 <- matrix(rnorm(d1$n_samples * 4), nrow = d1$n_samples)
desc2 <- desc1 + matrix(rnorm(length(desc1), sd = 0.03), nrow = nrow(desc1))
desc3 <- desc1 + matrix(rnorm(length(desc1), sd = 0.04), nrow = nrow(desc1))

net2 <- fm_network_match(
  domains = list(a = d1, b = d2, c = d3),
  descriptors = list(a = desc1, b = desc2, c = desc3),
  edges = c("a->b", "b->c", "c->a"),
  pair_batch_size = 2,
  descriptor_batch_size = 8,
  optimizer = "cg"
)
```
