---
title: "Functional Maps for Non-Shape Multidimensional Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functional Maps for Non-Shape Multidimensional Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This example treats each sample as a multidimensional observation vector rather than a mesh vertex. The two domains below are different feature-space views of the same latent samples, with a known row permutation.

```{r}
library(fmapsR)
set.seed(123)

make_view <- function(latent, weight, noise_sd = 0.02) {
  scale(latent %*% weight + matrix(rnorm(nrow(latent) * ncol(weight), sd = noise_sd), nrow(latent)))
}

make_multidomain <- function(X, k = 8) {
  operator <- fm_operator(X, method = "knn_laplacian", k = 7)

  fm_basis(
    fm_domain_generic(n_samples = nrow(X), data = X, operator = operator),
    k = k,
    solver = "base",
    cache = FALSE
  )
}

n <- 30
p <- 12
t <- seq(0, 2 * pi, length.out = n)
latent <- cbind(sin(t), cos(t), sin(2 * t), cos(2 * t))

W1 <- matrix(rnorm(ncol(latent) * p), ncol = p)
W2 <- W1 + 0.05 * matrix(rnorm(ncol(latent) * p), ncol = p)

truth <- c(4:n, 1:3)
X_A <- make_view(latent, W1)
X_B <- make_view(latent[truth, , drop = FALSE], W2)

A <- make_multidomain(X_A)
B <- make_multidomain(X_B)

stopifnot(
  all(dim(A$operator) == c(A$n_samples, A$n_samples)),
  max(abs(A$operator - t(A$operator))) < 1e-8
)

fit <- fm_match(A, B, descriptors = list(source = X_A, target = X_B), penalties = list(descr = 1, lap = 1e-3, comm = 0.2), optimizer = "cg", cg_maxit = 6)
fit <- fm_refine(fit, method = "icp", nit = 3)

metrics <- fm_fit_metrics(fit, truth = truth)
shared_signal <- latent[, 1] + 0.4 * latent[, 3]
target_signal <- shared_signal[truth]
signal_hat <- fm_transfer(fit, shared_signal)
baseline_mse <- mean((mean(target_signal) - target_signal)^2)

c(
  exact_match_rate = mean(as_p2p(fit) == truth),
  normalized_source_distance_error = metrics$geodesic$normalized_mean,
  transfer_mse = mean((signal_hat - target_signal)^2),
  baseline_mse = baseline_mse
)
```

## Takeaway

The API does not assume shapes. As long as you define:

1. Sample-wise descriptors (`n_samples x p`).
2. A basis-friendly operator (`n_samples x n_samples`).

you can run the same match/refine/transfer pipeline on tabular or embedding-style data.
