---
title: Pairwise Functional Maps
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2.0
    css: albers.css
    includes:
      in_header: albers-header.html
params:
  family: teal
  preset: homage
resource_files:
- albers.css
- albers.js
- albers-header.html

vignette: |
  %\VignetteIndexEntry{Pairwise Functional Maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(fmapsR)
set.seed(7)

make_path_adj <- function(n) {
  A <- matrix(0, nrow = n, ncol = n)
  A[cbind(1:(n - 1), 2:n)] <- 1
  A[cbind(2:n, 1:(n - 1))] <- 1
  A
}

make_pairwise_toy <- function(n = 24, shift = 2, k = 10, noise_sd = 0.005) {
  A <- make_path_adj(n)
  op <- diag(rowSums(A)) - A
  theta <- seq(0, 1, length.out = n)

  source_data <- cbind(
    theta,
    sin(2 * pi * theta),
    cos(2 * pi * theta),
    sin(4 * pi * theta),
    cos(6 * pi * theta)
  )

  truth <- c((shift + 1):n, seq_len(shift))
  target_data <- source_data[truth, , drop = FALSE] +
    matrix(rnorm(n * ncol(source_data), sd = noise_sd), nrow = n)

  source <- fm_basis(
    fm_domain_generic(
      n_samples = n,
      data = source_data,
      operator = op,
      adjacency = A
    ),
    k = k,
    solver = "base",
    cache = FALSE
  )

  target <- fm_basis(
    fm_domain_generic(
      n_samples = n,
      data = target_data,
      operator = op[truth, truth],
      adjacency = A[truth, truth]
    ),
    k = k,
    solver = "base",
    cache = FALSE
  )

  list(source = source, target = target, truth = truth)
}

toy <- make_pairwise_toy()
source <- toy$source
target <- toy$target
truth <- toy$truth

hks_source <- fm_descriptor_hks(source, n_times = 6)
hks_target <- fm_descriptor_hks(target, n_times = 6)

desc_source <- cbind(scale(source$data), hks_source)
desc_target <- cbind(scale(target$data), hks_target)
```

## End-to-end pairwise workflow

This toy pair has a real ground-truth correspondence (`truth`), so the metrics below are genuine validation rather than self-comparison.

```{r}
fit <- fm_match(source, target, descriptors = list(source = desc_source, target = desc_target), optimizer = "cg", cg_maxit = 6)
fit <- fm_refine(fit, method = "icp", nit = 5)
mapped_signal <- fm_transfer(fit, source$data[, 1])
p2p <- as_p2p(fit)
fm_fit_matrix(fit)[1:3, 1:3]
c(
  exact_match_rate = mean(p2p == truth),
  transfer_mse = mean((mapped_signal - target$data[, 1])^2),
  unique_source_matches = length(unique(p2p))
)
```

## Map Quality Metrics

`fm_fit_metrics()` provides coverage by default. Here the domains also carry graph adjacency and a known target-to-source correspondence, so we can report both normalized source-distance error and a continuity proxy.

```{r}
metrics <- fm_fit_metrics(fit, truth = truth)

c(
  coverage_ratio = metrics$coverage$ratio,
  normalized_geodesic = metrics$geodesic$normalized_mean,
  continuity_proxy = metrics$continuity$normalized_mean
)
```

## Why this interface is high-level

- The same `fm_domain_generic()` constructor can carry raw sample data, an operator, and adjacency for evaluation.
- Basis construction, matching, refinement, transfer, and pointwise conversion stay as separate steps.
- The quality helper can combine coverage, ground-truth distance error, and continuity when the inputs provide the needed structure.
