---
title: "Pairwise Functional Maps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pairwise Functional Maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(fmapsR)
set.seed(7)

make_domain <- function(phase, n = 24, k = 8) {
  t <- seq(0, 2 * pi, length.out = n)
  x <- cbind(sin(t + phase), cos(t + phase), sin(2 * t + phase))
  fm_basis(
    fm_domain_generic(n_samples = n, data = x, operator = diag(seq_len(n))),
    k = k,
    solver = "base",
    cache = FALSE
  )
}

source <- make_domain(phase = 0.00)
target <- make_domain(phase = 0.18)

# Basis-derived descriptor options
hks_source <- fm_descriptor_hks(source, n_times = 6)
hks_target <- fm_descriptor_hks(target, n_times = 6)

# You can also use WKS descriptors instead:
# wks_source <- fm_descriptor_wks(source, n_energies = 6)
# wks_target <- fm_descriptor_wks(target, n_energies = 6)

desc_source <- cbind(source$data, hks_source)
desc_target <- cbind(target$data, hks_target)
```

## End-to-end pairwise workflow

The core workflow below uses **6 lines** of API calls.

```{r}
fit <- fm_match(source, target, descriptors = list(source = desc_source, target = desc_target), optimizer = "cg", cg_maxit = 4)
fit <- fm_refine(fit, method = "icp", nit = 3)
mapped_signal <- fm_transfer(fit, source$data[, 1])
p2p <- as_p2p(fit)
fm_fit_matrix(fit)[1:3, 1:3]
c(mse = mean((mapped_signal - target$data[, 1])^2), p2p_count = length(p2p))
```

## Map Quality Metrics

`fm_fit_metrics()` provides coverage by default and adds continuity/geodesic-style scores when adjacency or ground truth are available.

```{r}
metrics <- fm_fit_metrics(fit, truth = p2p)

c(
  coverage_ratio = metrics$coverage$ratio,
  geodesic_mean = metrics$geodesic$mean,
  continuity_mean = metrics$continuity$mean
)
```

## Why this interface is high-level

- Domains use one constructor (`fm_domain*`) regardless of geometry type.
- Basis + fit + refine are explicit and composable.
- Transfer and pointwise conversion are one-liners.
