---
title: Shape Correspondence on Posed Meshes
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2.0
    css: albers.css
    includes:
      in_header: albers-header.html
params:
  family: teal
  preset: homage
resource_files:
- albers.css
- albers.js
- albers-header.html
- tosca-cats

vignette: |
  %\VignetteIndexEntry{Shape Correspondence on Posed Meshes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette has two goals.

First, it shows real benchmark meshes from the classic TOSCA cats used in early functional-maps papers. That gives the package a genuine shape-correspondence example instead of a matrix-only toy.

Second, it keeps one small bundled mesh pair with known exact correspondence so the package can still demonstrate a fully accurate end-to-end workflow in a form that is cheap to render and regression-test.

## Start with the real TOSCA cats

```{r load-real-cats}
library(fmapsR)

pair_real <- fm_example_mesh_pair("tosca_cat10")
```

`fm_example_mesh_pair("tosca_cat10")` loads vendored meshes from the TOSCA high-resolution cat benchmark:

- `cat0` is the source shape
- `cat10` is the target pose
- the original dataset gives a known vertex correspondence through shared vertex ordering
- bundled thumbnails make it easy to show the original benchmark poses directly in the article

```{r show-tosca-gallery, echo = FALSE, out.width = "19%"}
knitr::include_graphics(file.path(
  "tosca-cats",
  c("cat0.png", "cat1.png", "cat2.png", "cat6.png", "cat10.png")
))
```

The thumbnails above are the original TOSCA pose renders for `cat0`, `cat1`, `cat2`, `cat6`, and `cat10`. They are the same family of meshes referenced in the Ovsjanikov et al. functional-maps paper.

## Inspect the actual benchmark pair

```{r real-geometry-helpers, include = FALSE}
render_real_geometry_pair <- function() {
  old_par <- graphics::par(no.readonly = TRUE)
  on.exit(graphics::par(old_par), add = TRUE)
  graphics::par(mfrow = c(1, 2), mar = c(1, 1, 3, 1))

  neutral_palette <- c("#d6d3d1", "#d6d3d1")
  neutral_values <- rep(1, nrow(pair_real$source$vertices))

  fm_plot_mesh_scalar(
    pair_real$source,
    neutral_values,
    engine = "base",
    title = "cat0 source pose",
    palette = neutral_palette,
    value_range = c(0, 1),
    view = "yz"
  )
  fm_plot_mesh_scalar(
    pair_real$target,
    neutral_values,
    engine = "base",
    title = "cat10 target pose",
    palette = neutral_palette,
    value_range = c(0, 1),
    view = "yz"
  )
}
```

```{r show-real-geometry, echo = FALSE, fig.width = 9.2, fig.height = 3.5}
render_real_geometry_pair()
```

This is the actual mesh pair used below, rendered directly from the bundled vertices and faces rather than from the original dataset thumbnails.

## Build a real cat-to-cat map

For the real cats, this example uses a cotangent mesh operator with landmark-HKS descriptors and a short ICP refinement pass. The landmark indices are known correspondences supplied to both shapes, so this is a supervised demonstration. The transfer scores below describe this pair and signal; they do not establish unsupervised matching accuracy or performance across TOSCA.

```{r fit-real-cats}
op_source <- fm_operator_mesh(
  pair_real$source$vertices,
  pair_real$source$faces,
  method = "cotangent"
)
op_target <- fm_operator_mesh(
  pair_real$target$vertices,
  pair_real$target$faces,
  method = "cotangent"
)

source_real <- fm_basis(
  fm_domain_mesh(pair_real$source$vertices, pair_real$source$faces, operator = op_source),
  k = 30,
  solver = "rspectra",
  cache = FALSE
)
target_real <- fm_basis(
  fm_domain_mesh(pair_real$target$vertices, pair_real$target$faces, operator = op_target),
  k = 30,
  solver = "rspectra",
  cache = FALSE
)

desc_source_real <- fm_descriptor_hks(source_real, n_times = 6, landmarks = pair_real$landmarks)
desc_target_real <- fm_descriptor_hks(target_real, n_times = 6, landmarks = pair_real$landmarks)

fit_real <- fm_match(
  source_real,
  target_real,
  descriptors = list(source = desc_source_real, target = desc_target_real),
  penalties = list(descr = 1, lap = 1e-4, comm = 0.02),
  init = "identity",
  optimizer = "cg",
  cg_maxit = 300,
  cg_tol = 1e-8
)
stopifnot(fit_real$diagnostics$convergence == 0L)
fit_real <- fm_refine(fit_real, method = "icp", nit = 1)
```

## Compare ground-truth and estimated heat transfer

```{r real-transfer-helpers, include = FALSE}
source_center <- pair_real$source$vertices[pair_real$landmarks[[5]], ]
source_signal_real <- exp(-rowSums(sweep(pair_real$source$vertices, 2, source_center, "-")^2) / (2 * 20^2))
target_signal_real <- source_signal_real[pair_real$truth]
signal_hat_real <- fm_transfer(fit_real, source_signal_real)
signal_error_real <- abs(signal_hat_real - target_signal_real)

real_top_overlap <- function(frac) {
  top_n <- max(50L, floor(length(target_signal_real) * frac))
  length(intersect(
    order(signal_hat_real, decreasing = TRUE)[seq_len(top_n)],
    order(target_signal_real, decreasing = TRUE)[seq_len(top_n)]
  )) / top_n
}

real_transfer_stats <- list(
  correlation = stats::cor(signal_hat_real, target_signal_real),
  overlap_1 = real_top_overlap(0.01),
  overlap_5 = real_top_overlap(0.05),
  overlap_10 = real_top_overlap(0.10),
  mse = mean((signal_hat_real - target_signal_real)^2)
)

stopifnot(
  is.finite(real_transfer_stats$correlation),
  real_transfer_stats$correlation >= 0.95,
  real_transfer_stats$overlap_5 >= 0.95,
  real_transfer_stats$overlap_10 >= 0.90,
  real_transfer_stats$mse <= 0.005
)

real_metrics <- data.frame(
  metric = c(
    "Transfer correlation",
    "Top-1% support overlap",
    "Top-5% support overlap",
    "Top-10% support overlap",
    "Transfer MSE"
  ),
  value = c(
    real_transfer_stats$correlation,
    real_transfer_stats$overlap_1,
    real_transfer_stats$overlap_5,
    real_transfer_stats$overlap_10,
    real_transfer_stats$mse
  )
)

render_real_transfer_panels <- function() {
  old_par <- graphics::par(no.readonly = TRUE)
  on.exit(graphics::par(old_par), add = TRUE)
  graphics::par(mfrow = c(1, 4), mar = c(1, 1, 3, 1))
  value_range <- range(c(source_signal_real, target_signal_real, signal_hat_real), finite = TRUE)

  fm_plot_mesh_scalar(
    pair_real$source,
    source_signal_real,
    engine = "base",
    title = "cat0 source heat",
    show_edges = FALSE,
    view = "yz",
    value_range = value_range
  )
  fm_plot_mesh_scalar(
    pair_real$target,
    target_signal_real,
    engine = "base",
    title = "cat10 ground truth",
    show_edges = FALSE,
    view = "yz",
    value_range = value_range
  )
  fm_plot_mesh_scalar(
    pair_real$target,
    signal_hat_real,
    engine = "base",
    title = "cat10 estimated transfer",
    show_edges = FALSE,
    view = "yz",
    value_range = value_range
  )
  fm_plot_mesh_scalar(
    pair_real$target,
    signal_error_real,
    engine = "base",
    title = "Absolute error",
    show_edges = FALSE,
    view = "yz",
    palette = c("#f8fafc", "#f59e0b", "#b91c1c"),
    value_range = range(signal_error_real, finite = TRUE)
  )
}
```

```{r show-real-transfer, echo = FALSE, fig.width = 13.5, fig.height = 3.6}
render_real_transfer_panels()
```

The second panel is the exact target heatmap implied by the dataset’s known correspondence. The third panel is what `fmapsR` estimates from the source signal alone. On this `cat0 -> cat10` pair the transferred field stays on the correct head-and-shoulder region, tracks the body attenuation well, and broadly preserves the target pose.

```{r show-real-metrics}
knitr::kable(real_metrics, digits = 4)
```

This is still not a paper-perfect pointwise correspondence result. Exact nearest-vertex recovery remains much weaker than the smooth transfer shown above, so the right success criterion here is: does the spectral map move a localized function to the correct target region with low distortion? On that criterion, the current cat example is finally strong enough to teach from.

## Keep one exact small-mesh sandbox

The bundled toy pair checks an intentionally easy case: both domains use the same topology-derived operator and corresponding synthetic descriptors. Its exact correspondence scores are a workflow regression check, not evidence of accuracy on independently sampled or unfamiliar shapes.

```{r load-toy}
pair_toy <- fm_example_mesh_pair("posed_tube")
op_toy <- fm_operator_mesh(
  pair_toy$source$vertices,
  pair_toy$source$faces,
  method = "face_graph",
  weight_mode = "binary"
)

source_toy <- fm_basis(
  fm_domain_mesh(pair_toy$source$vertices, pair_toy$source$faces, operator = op_toy),
  k = 24,
  solver = "rspectra",
  cache = FALSE
)
target_toy <- fm_basis(
  fm_domain_mesh(pair_toy$target$vertices, pair_toy$target$faces, operator = op_toy),
  k = 24,
  solver = "rspectra",
  cache = FALSE
)

desc_source_toy <- cbind(
  scale(pair_toy$source$descriptors),
  fm_descriptor_hks(source_toy, n_times = 8, landmarks = pair_toy$landmarks)
)
desc_target_toy <- cbind(
  scale(pair_toy$target$descriptors),
  fm_descriptor_hks(target_toy, n_times = 8, landmarks = pair_toy$landmarks)
)

fit_toy <- fm_match(
  source_toy,
  target_toy,
  descriptors = list(source = desc_source_toy, target = desc_target_toy),
  penalties = list(descr = 1, lap = 1e-4, comm = 0.01),
  init = "identity",
  optimizer = "cg",
  cg_maxit = 8
)
fit_toy <- fm_refine(fit_toy, method = "icp", nit = 4)
```

```{r toy-transfer-helpers, include = FALSE}
toy_center <- pair_toy$source$vertices[pair_toy$landmarks[[3]], ]
toy_signal <- exp(-rowSums(sweep(pair_toy$source$vertices, 2, toy_center, "-")^2) / (2 * 0.18^2))
toy_signal_hat <- fm_transfer(fit_toy, toy_signal)
toy_metrics <- fm_fit_metrics(fit_toy, truth = pair_toy$truth)

toy_table <- data.frame(
  metric = c(
    "Exact match rate",
    "Coverage ratio",
    "Normalized geodesic error",
    "Transfer MSE"
  ),
  value = c(
    mean(as_p2p(fit_toy) == pair_toy$truth),
    toy_metrics$coverage$ratio,
    toy_metrics$geodesic$normalized_mean,
    mean((toy_signal_hat - toy_signal[pair_toy$truth])^2)
  )
)

render_toy_transfer_panels <- function() {
  old_par <- graphics::par(no.readonly = TRUE)
  on.exit(graphics::par(old_par), add = TRUE)
  graphics::par(mfrow = c(1, 3), mar = c(1, 1, 3, 1))

  fm_plot_mesh_scalar(pair_toy$source, toy_signal, engine = "base", title = "Toy source heat", show_edges = TRUE)
  fm_plot_mesh_scalar(pair_toy$target, toy_signal[pair_toy$truth], engine = "base", title = "Toy target truth", show_edges = TRUE)
  fm_plot_mesh_scalar(pair_toy$target, toy_signal_hat, engine = "base", title = "Toy estimated transfer", show_edges = TRUE)
}
```

```{r show-toy-transfer, echo = FALSE, fig.width = 9.5, fig.height = 3.4}
render_toy_transfer_panels()
```

```{r show-toy-metrics}
knitr::kable(toy_table, digits = 4)
```

That split is deliberate:

- the real TOSCA cats provide the recognizable benchmark visuals
- the small bundled mesh pair provides the exact quantitative workflow that the package can currently defend end to end
